---
title: "GV300 Showcase"
output: distill::distill_article
---
# Introduction
This work demonstrates some of the work I have undertaken as part of my
undergraduate studies. It explores the use of quantitative methods on text,
i.e. text analysis.

The first section investigates elements of textual data, specifically whether
UK political party manifestos are getting shorter or longer. The first graph
shows that over time, party manifestos have lengthened, demonstrated via the
rise in the number of sentences found in each manifesto. The second graph seeks
to understand this relationship further by splitting the data by party to
understand if this trend is observed across all parties. It finds that it is.

The second section conducts a simple Naive Bayes classifier which is a statistical
classification technique that can be used for text classification. In this case
it classifies text into Republican or Democrat. The classifier is found to 
perform poorly in this context.

# The rise in the number of sentences in political manifestos

*Calculating sentence numbers*
```{r, sentence numbers}
library(readtext)
library(quanteda)
library(ggplot2)

manifestos <- readtext("UK Manifestos/*.txt", docvarsfrom = c("filenames"))

manifestos_corpus <- corpus(manifestos, docnames = manifestos$doc_id)

num_sentences <- nsentence(manifestos_corpus)
head(num_sentences)

```
This section loads the data, turns the text files into a 'corpus' and shows how
many sentences are in each manifesto.


*Sorting dates*
```{r, Dates}

nms <- names (manifestos_corpus)

dates <- as.numeric(gsub('[^[:digit:]]','', nms))
head(dates)

```
This has created new objects for the names of the text files and the dates
each manifesto was published in.


*Plotting the number of sentences by year*
```{r, Sentence plot}

sentence_df <- data.frame(num_sentences, dates)

sentence_df$doc_id <- manifestos$doc_id

sentence_df$party <- substr(sentence_df$doc_id, 1, 3)
  
ggplot(sentence_df, aes(x = dates, y = num_sentences))+
  geom_point()+
  labs(
  x = "Year",
  y = "Number of Sentences") +
  theme(
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black",size = 1, fill = NA),
    panel.background = element_rect(fill = "white"))

```
A new data frame was created for the graph. 'doc_id' is added to the data frame 
for the next graph that explores whether all parties see an increase in 
manifesto length. The separate parties are also identified.

The graph shows that there has been a increase in the number of sentences in 
manifestos over the years.


*Plotting the number of sentences by party*
```{r, Party investigation}

ggplot(sentence_df, aes(x = dates, y = num_sentences))+
  geom_point()+
  facet_wrap(~party)+
  labs(
  x = "Year",
  y = "Number of Sentences") +
  theme(
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black",size = 1, fill = NA),
    panel.background = element_rect(fill = "white"))

```
This second graph shows that the growth of manifest length is consistent across 
all parties.


# Naive Bayes Classification

*Developing a detailed data set*
```{r, Speech data}

library(stringr)
library(quanteda.textmodels)
library(lattice)
library(caret)

speeches <- readtext("Concession Speeches/*.txt") 

speeches_corpus <- corpus(speeches)

nam <-names(speeches_corpus)

speeches_corpus$Dates  <- paste(as.numeric(gsub('[^[:digit:]]','',
                                                names(speeches_corpus))))

speeches_corpus$Party <- str_sub(nam, start= -5, -5)

speeches_corpus$PNames <- str_sub(nam, end = -12)

summary(speeches_corpus)

```
This firstly loads the data and creates a 'corpus'. It collects the names, dates,
parties, and president names. The summary shows the new columns which provides
useable data for analysis and classification.


*Creating paragraphs*
```{r, Creating paragraphs}

speeches_paragraph <- corpus_reshape(speeches_corpus, to = "paragraphs", 
                                     use_docvars = TRUE)

table(speeches_paragraph$Party)

```
To increase the data available, the speeches are split into paragraphs. I check
the data to see if the balance of political parties is equal because an imbalance
will see the classifier over predict. In this case, there is an imbalance with
more Republicans than Democrats.


*Data cleaning*
```{r, Data cleaning}

speech_tokens <- tokens(speeches_paragraph, remove_punct = TRUE,
                        remove_symbols = TRUE, split_hyphens = TRUE,
                        remove_numbers = TRUE)

speeches_clean <- tokens_wordstem(speech_tokens)

speeches_clean <- tokens_remove(speeches_clean,
                                pattern = c("applpaus", "applause"))

speeches_clean <- tokens_select(speeches_clean, c("*st", "*th", "*rd"),
                                selection = "remove")

speeches_clean <- dfm(speeches_clean)

speeches_clean <- dfm_remove(speeches_clean,
                      stopwords("english"))

speeches_clean <- speeches_clean[, order(featnames(speeches_clean))]

```
The first line of code provides a list of words from the speeches. Then the
ending of words are removed, spelling mistakes are removed, and special endings
such as the 'st' in '1st' are removed. Then a data frame matrix is made which
provides a better structure for analysis. Filler words such as 'and' and 'the'
are removed and the remaining words are ordered alphabetically.


*Training the system*
```{r, Training}

Concession_dfm_train <- speeches_clean[1:268, ]

Concession_dfm_test <- speeches_clean[269:337, ]

Concession_train_labels <- speeches_paragraph$Party[1:268]

Concession_test_labels <- speeches_paragraph$Party[269:337]

Concession_dfm_train

Concession_train_labels[1:10]

```
This chunk trains the system using an 80/20 split, meaning that the model is
trained on 80% of the data and the remaining 20% is used to evaluate its 
performance.


*Applying the text model*
```{r, Model}

nb_model <- textmodel_nb(Concession_dfm_train, docvars(Concession_dfm_train,
                                                       "Party"))
summary(nb_model)


```
The model assumes that all words are independent from one another (doesn't
consider how they function in a sentence) and that their position doesn't matter.
To understand if the results are accurate further evaluation is needed.


*Removing words not included in the training vocabulary*
```{r, Removing words}

dfmat_matched <- dfm_match(Concession_dfm_test, features =
                             featnames(Concession_dfm_train))
                           
actual_class <- dfmat_matched$Party

predicted_class <- predict(nb_model, newdata = dfmat_matched)
```
Because some words in the test data were not included in the training data they
are removed and no probability is included in the model.


*Checking if the system works*
```{r, matrix}

tab_class <- table(predicted_class, actual_class)

cfm <- confusionMatrix(tab_class, mode="everything")
cfm

```
This creates a confusion matrix which visualises how the algorithm regarding the
human 'gold-standard':

top left: true positive
bottom left: false negative

top right: false negative
bottom right: true negative

This feedback shows that the accuracy of the system is 59%. The recall of 0.83
shows that the system finds most of the Democrat cases however, with a precision of 0.2,
it is usually wrong. Ultimately, the model is unreliable.


# Link to repository
https://github.com/martha-png/GV300-Assignment
You can also embed plots, for example:
